---
title: "curve_stripping"
author: "Devin"
date: "August 19, 2014"
output: html_document
---


```{r interactive, include=FALSE}
setwd(paste0(getwd(), "/lab-notebook"))
```


```{r}
library(dplyr)
library(ggplot2)
```

```{r}
dat <- read.table("../modeling//singleid_sim.dat", skip = 1, header = T)

```

```{r}
qplot(data=dat, x = TIME, y = PRED) + scale_y_log10()
```


## Curve stripping flow

* get cmax and filter all times before
* fit linear model through log transformed data to get terminal slope
* subtract predicted values from terminal portion from observed values to get portion remainder that corresponds to distribution portion
* fit second line through the part corresponding to distribution portion to calculate rate constant alpha

* convert A, B $\alpha$ and $\beta$ to physiologic parameter versions


```{r}
time<- dat$TIME
  conc <- dat$DV
  
  time.points <- length(time)
  #check to make sure partial time legit option
  ###need to add warning
last_points <- 5
  last <- last_points
  start<- time.points - last + 1


    t<-time[start:time.points]
    con <- conc[start:time.points]
    xt <-lm(log(con)~t) # log-linear terminal phase calculation for k
    lambda_z<- as.numeric(xt$coef[2])

predicted_terminal_portion <- exp(predict(xt, data.frame(t = time)))
remaining_alpha_conc <- conc- predicted_terminal_portion

alpha_index <- 4:9
alpha_conc <- remaining_alpha_conc[alpha_index]
alpha_t <- time[alpha_index]
alpha_xt <- lm(log(alpha_conc)~alpha_t)

A <- exp(unlist(alpha_xt$coef[1]))
B <- exp(unlist(xt$coef[1]))
alpha <- -as.numeric(unlist(alpha_xt$coef[2]))
beta <- -as.numeric(unlist(xt$coef[2]))
```

for calculations see boomer

http://www.boomer.org/c/p4/c19/c19.pdf

```{r}
DOSE <- 1000
k21 <- (A*beta + B*alpha)/(A+B)
kel <- alpha*beta/k21
k12 <- alpha + beta - k21 - kel

Vc <- DOSE/max(conc)
Vp <-  k12/k21*Vc
Q <- k12*Vc
CL <- kel*Vc

initial_estimates <- data.frame("Vc" = Vc, "Vp" = Vp, "Q" = Q, "CL" = CL)
row.names(initial_estimates) <- NULL
round(initial_estimates, 2)
dat %>% do(data.frame(AUC_inf(.$TIME, .$DV, AUCinf_only = F)))
```

```{r}
strip_curves <- function(TIME, DV, DOSE, number_terminal_points, oral= FALSE) {
  time<- TIME
  conc <- DV
  
  num_time_points <- length(time)
  #check to make sure partial time legit option
  ###need to add warning
start_terminal_points <- num_time_points - number_terminal_points + 1


    terminal_time<-time[start_terminal_points:num_time_points]
    terminal_conc <- conc[start_terminal_points:num_time_points]
    terminal_xt <-lm(log(terminal_conc)~terminal_time) # log-linear terminal phase calculation for k
    terminal_lambda_z<- as.numeric(terminal_xt$coef[2])

predicted_terminal_portion <- exp(predict(terminal_xt, data.frame(terminal_time = time)))
remaining_alpha_conc <- conc- predicted_terminal_portion

alpha_index <- 1:(num_time_points - number_terminal_points)
if(oral) {
  # cmax to last point not in terminal phase 
  alpha_index <- which(conc==max(conc)):(num_time_points - number_terminal_points)
}
alpha_conc <- remaining_alpha_conc[alpha_index]
alpha_t <- time[alpha_index]
alpha_xt <- lm(log(alpha_conc)~alpha_t)

A <- exp(unlist(alpha_xt$coef[1]))
B <- exp(unlist(terminal_xt$coef[1]))
alpha <- -as.numeric(unlist(alpha_xt$coef[2]))
beta <- -as.numeric(unlist(terminal_xt$coef[2]))

k21 <- (A*beta + B*alpha)/(A+B)
kel <- alpha*beta/k21
k12 <- alpha + beta - k21 - kel

Vc <- DOSE/max(conc)
Vp <-  k12/k21*Vc
Q <- k12*Vc
CL <- kel*Vc

initial_estimates <- data.frame("Vc" = Vc, "Vp" = Vp, "Q" = Q, "CL" = CL)
row.names(initial_estimates) <- NULL
return(round(initial_estimates, 2))
#dat %>% do(data.frame(AUC_inf(.$TIME, .$DV, AUCinf_only = F)))
}

strip_curves(dat$TIME, dat$DV, DOSE = 1000, number_terminal_points=5, oral=T)
```





